Queue Aware Precoder Design for Space Frequency Resource Allocation

### 2014-07-11 Report/dual_approach.tex
Updated MSE KKT

In contrast to the \acl{PD} problem, the \acf{DD} performs the distributed precoder design by relaxing the interference constraints by including it in the objective function of each subproblem with a penalty pricing \cite{tolli2011decentralized,palomar2006tutorial}. In order to decouple the problem \eqref{eqn-decent-1}, the global coupling variables \me{\zeta_{l,k,n,b}} in \eqref{eqn-decent-3} are replaced by the local copies at each \ac{BS} \me{b} denoted by \me{\zeta^{\set{b}}_{l,k,n,b}, \fall b\in \mc{B}}, which is then used in the problem as an optimization variable.

Let  \me{\mbfa{\zeta}^{\set{b_k}}} be the locally maintained vector formed by stacking the interference entries relevant to the \ac{BS} \me{b_k}. Let \me{\mbfa{\zeta}} be the global interference entries of all \acp{BS} in the set \me{\mc{B}} stacked as a vector as
\begin{IEEEeqnarray}{rCl}
\mbfa{\zeta} &=& \left [ \zeta_{1,\bar{\mc{U}}_{1}(1),1,1}, \dotsc, \zeta_{L,\bar{\mc{U}}_{1}(1),1,1}, \dotsc, \zeta_{L,\bar{\mc{U}}_{1}(|\bar{\mc{U}}_{1}|),1,1},  \right. \nonumber \\
&\quad& \left . \dotsc, \zeta_{L,\bar{\mc{U}}_{N_B}(|\bar{\mc{U}}_{N_B}|),1,N_B}, \dotsc, \zeta_{L,\bar{\mc{U}}_{N_B}(|\bar{\mc{U}}_{N_B}|),N,N_B} \right ] \IEEEyessubnumber \\
n_{b_k} &=& |\mbfa{\zeta}^{\set{b_k}}| = N L \sum_{b \in \mc{B}} \left | \bar{\mc{U}}_b\right |, \IEEEyessubnumber
\end{IEEEeqnarray}
where \me{\mbfa{\zeta}(b_k)} denote the entries corresponding to the \ac{BS} \me{b_k} and the vector \me{\mbfa{\nu}^{\set{b_k}}} stacks the dual variables corresponding to the equality condition \me{\mbfa{\zeta}^{\set{b_k}} = \mbfa{\zeta}(b_k)}. The equality constraint for the local and the global interference vector \me{\zeta^{\set{b_k}}_{l,k,n,b} = \zeta_{l,k,n,b}, \fall b \in \bar{\mc{B}}_{b_k}, k \in \mc{U}} and \me{\forall k \in \bar{\mc{U}}_{b_k}, b=b_k} is relaxed by the partial Lagrangian in the objective as
\begin{IEEEeqnarray}{rCl} \label{eqn-dual-2} \label{eqn-dual-1}
\underset{\substack{\gamma_{l,k,n}, \nu^{\set{b_k}}_{l,k,n,b} \\ \mvec{m}{l,k,n}, \beta_{l,k,n}, \zeta^{\set{b_k}}_{l,k,n,b}}}{\text{minimize}} &\quad & \| \tilde{\mbf{v}}_{b_k} \|_q + \mbfa{\nu}^{{\set{b_k}}T} \left ( \mbfa{\zeta}^{\set{b_k}} - \mbfa{\zeta}(b_k) \right ) \IEEEyessubnumber \label{eqn-dual-2a} \\
\text{subject to} && \beta_{l,k,n} \geq \sum_{\substack{j = 1\\j \neq l}}^L |\mvec{w}{l,k,n}^\herm \mvec{H}{{b_k},k,n} \mvec{m}{j,k,n} |^2 \nonumber \\
&&\quad + \sum_{i \in \mc{U}_{b_k} \backslash \{k\}} \sum_{j = 1}^L |\mvec{w}{l,k,n}^\herm \mvec{H}{{b_k},k,n} \mvec{m}{j,i,n} |^2 + \sum_{b \in \bar{\mc{B}}_{b_k}} \zeta^{\set{b_k}}_{l,k,n,b} \; + \; N_0 \IEEEyessubnumber \label{eqn-dual-1d} \\
&& \zeta^{\set{b_k}}_{\pr{l},\pr{k},n,{b_k}} \geq \sum_{k \in \mc{U}_b} \sum_{l = 1}^L |\mvec{w}{\pr{l},\pr{k},n}^\herm \mvec{H}{b_k,\pr{k},n} \mvec{m}{l,k,n} |^2, \; \forall \pr{k} \in \bar{\mc{U}}_{b_k}, \; \forall n \in \mc{C} \IEEEyessubnumber \label{eqn-dual-1e} \\
&\quad& \mbfa{\nu}^{\set{b_k}} \in \mathbb{R}^{n_{b_k}}_+, \; \eqref{eqn-8} \; \text{and} \; \eqref{eqn-primal-1c}. \IEEEyessubnumber \label{eqn-dual-1f}
\end{IEEEeqnarray}
In \eqref{eqn-dual-2a}, the global interference variable \me{\mbfa{\zeta}(b_k)} is dropped without affecting the optimal solution, since it is fixed for a given iteration. The problem defined in \eqref{eqn-dual-2} can be decoupled to solve for the precoders at each \ac{BS} by using interference vector \me{\mbfa{\zeta}^{\set{b_k}}} as an optimization variable for the fixed interference price (dual variable) \me{\mbfa{\nu}^{\set{b_k}}}. In the current work, we consider the robust counterpart of the \acl{DD} scheme, namely \ac{ADMM} approach, which doesn't require any assumptions on the convexity of the functions for convergence \cite{boyd2011distributed}.

In order to bound the interference assumptions \me{\zeta^{\set{b_k}}_{l,k,n,b}} and \me{\zeta^{\set{b}}_{l,k,n,b}} between the \acp{BS} \me{b_k} and \me{b}, the \ac{ADMM} method augments a scaled quadratic penalty of the interference deviation to the objective function \cite{bertsekas1999nonlinear,boyd2011distributed}. At the optimal point, the actual and the assumed (local) interference values will be equal and provides no contribution to the objective function. It can be shown that the optimal step size is equal to the scaling factor \me{\rho} used for the penalty term in the objective function \cite{bertsekas1999nonlinear,boyd2011distributed}. Now, the subproblem at each \ac{BS} for the \me{\ith{i}} iteration is give by
\begin{IEEEeqnarray}{rCl} \label{eqn-dual-3}
\underset{\substack{\gamma_{l,k,n}, \mvec{m}{l,k,n}, \\ \beta_{l,k,n}, \zeta^{\set{b_k}(i)}_{l,k,n,b}}}{\text{minimize}} &\quad & \| \tilde{\mbf{v}}_{b_k} \|_q + \mbfa{\nu}^{{\set{b_k}(i)}T} \mbfa{\zeta}^{\set{b_k}(i)} + \frac{\rho}{2} \Big \| \, \underbrace{\mbfa{\zeta}^{(i)}(b_k)}_{\text{global iterate}} - \underbrace{\mbfa{\zeta}^{\set{b_k}(i)}}_{\text{local iterate}} \, \Big \|^2_2 \IEEEyessubnumber \label{eqn-dual-3a} \\
\text{subject to} & \quad & \eqref{eqn-dual-1d} - \eqref{eqn-dual-1f} \IEEEyessubnumber \label{eqn-dual-3b},
\end{IEEEeqnarray}
where \me{(i)} represents the current iteration or the information exchange index and \me{\mbfa{\zeta}^{(i)}} denotes the updated global interference level from the \me{\ith{(i-1)}} information exchange of the local interference vector \me{\mbfa{\zeta}^{\set{b}(i-1)}, \forall b \in \mc{B}}.

Once the local problems are solved at each \ac{BS}, the update for the global interference vector \me{\mbfa{\zeta}^{(i)}} and the dual variables \me{\mbfa{\nu}^{\set{b}(i)}} can be performed at each \ac{BS} independently by exchanging the corresponding local copies of the interference vector \me{\mbfa{\zeta}^{\set{b}}, \forall b \in \mc{B}}. Since the entries in \me{\mbfa{\zeta}^{(i)}} relates exactly two \acp{BS} only, each entry in the \me{\mbfa{\zeta}^{(i)}} can be updated by exchanging the local copies from the corresponding two \acp{BS} only. For instance, the entry \me{\zeta^{(i)}_{l,\mc{U}_{b_k}(1),n,b}} depends on the local interference value \me{\zeta^{\set{b_k}(i-1)}_{l,\mc{U}_{b_k}(1),n,b}} assumed by the \ac{BS} \me{b_k} and the actual interference caused by the \ac{BS} \me{b} as in \me{\zeta^{\set{b}(i-1)}_{l,\mc{U}_{b_k}(1),n,b}} as
\begin{equation}
\zeta_{l,\mc{U}_{b_k}(1),n,b}^{(i)} = \frac{1}{2} \, \left ( \, \zeta^{\set{b}(i-1)}_{l,\mc{U}_{b_k}(1),n,b} + \zeta^{\set{b_k}(i-1)}_{l,\mc{U}_{b_k}(1),n,b} \, \right ).
\label{zeta_update}
\end{equation}
The dual variable entries in the vector \me{\mbfa{\nu}^{\set{b_k}}}, which is the stacked dual variables corresponding to the interference equality constraint at the \ac{BS} \me{b_k}, are updated using the subgradient as
\begin{equation}\label{dual-sg-update}
\nu^{\set{b_k}(i)}_{l,k,n,b} = \nu^{\set{b_k}(i-1)}_{l,k,n,b} + \rho \, \left (\zeta^{\set{b_k}(i-1)}_{l,k,n,b} - \zeta^{\set{b}(i-1)}_{l,k,n,b} \right ), \forall b,b_k \in \mc{B}, \forall k \in \bar{\mc{U}}_{b}.
\end{equation}

>>>>

In contrast to the \acl{PD} problem, the \acf{DD} relaxes the interference constraints by including it in the objective function of each subproblem with a penalty pricing \cite{palomar2006tutorial,boyd2011distributed}. In order to decouple the problem \eqref{eqn-decent-1}, the coupling variables \me{\zeta_{l,k,n,b}} in \eqref{eqn-decent-3} are replaced by the respective local copies \me{\mbfa{\zeta}^{\set{b}}, \forall b \in \mc{B}}, which are then solved for an optimal solution. Now the sub problems are coupled by the global consensus vector \me{\mbfa{\zeta}} maintaining the complete stacked  interference profile of all users in the system as
\begin{IEEEeqnarray}{rCl}
\mbfa{\zeta} &=& \left [ \zeta_{1,\bar{\mc{U}}_{1}(1),1,1}, \dotsc, \zeta_{L,\bar{\mc{U}}_{1}(1),1,1}, \dotsc, \zeta_{L,\bar{\mc{U}}_{1}(|\bar{\mc{U}}_{1}|),1,1},  \right. \nonumber \\
&\quad& \left . \dotsc, \zeta_{L,\bar{\mc{U}}_{N_B}(|\bar{\mc{U}}_{N_B}|),1,N_B}, \dotsc, \zeta_{L,\bar{\mc{U}}_{N_B}(|\bar{\mc{U}}_{N_B}|),N,N_B} \right ] \IEEEyessubnumber \\
n_{b_k} &=& |\mbfa{\zeta}^{\set{b_k}}| = N L \sum_{b \in \mc{B}} \left | \bar{\mc{U}}_b\right |. \IEEEyessubnumber
\end{IEEEeqnarray}

Let \me{\mbfa{\zeta}(b_k)} denotes the consensus entries corresponding to the \ac{BS} \me{b_k}. Let \me{\mbfa{\nu}^{\set{b_k}}} represents the stacked dual variables corresponding to the equality condition \me{\mbfa{\zeta}^{\set{b_k}} = \mbfa{\zeta}(b_k)} used in the subproblems. The equality constraint for the local and the consensus interference vector \me{\mbfa{\zeta}^{\set{b_k}} = \mbfa{\zeta}(b_k)} present in each subproblem is relaxed by the taking the partial Lagrangian for the constraint as
\begin{IEEEeqnarray}{rCl} \label{eqn-dual-2} \label{eqn-dual-1}
\underset{\substack{\gamma_{l,k,n}, \nu^{\set{b_k}}_{l,k,n,b} \\ \mvec{m}{l,k,n}, \beta_{l,k,n}, \zeta^{\set{b_k}}_{l,k,n,b}}}{\text{minimize}} &\quad & \| \tilde{\mbf{v}}_{b_k} \|_q + \mbfa{\nu}^{{\set{b_k}}T} \left ( \mbfa{\zeta}^{\set{b_k}} - \mbfa{\zeta}(b_k) \right ) \IEEEyessubnumber \label{eqn-dual-2a} \\
\text{subject to} && \beta_{l,k,n} \geq \sum_{\substack{j = 1\\j \neq l}}^L |\mvec{w}{l,k,n}^\herm \mvec{H}{{b_k},k,n} \mvec{m}{j,k,n} |^2 \nonumber \\
&&\quad + \sum_{i \in \mc{U}_{b_k} \backslash \{k\}} \sum_{j = 1}^L |\mvec{w}{l,k,n}^\herm \mvec{H}{{b_k},k,n} \mvec{m}{j,i,n} |^2 + \sum_{b \in \bar{\mc{B}}_{b_k}} \zeta^{\set{b_k}}_{l,k,n,b} \; + \; N_0 \IEEEyessubnumber \label{eqn-dual-1d} \\
&& \zeta^{\set{b_k}}_{\pr{l},\pr{k},n,{b_k}} \geq \sum_{k \in \mc{U}_b} \sum_{l = 1}^L |\mvec{w}{\pr{l},\pr{k},n}^\herm \mvec{H}{b_k,\pr{k},n} \mvec{m}{l,k,n} |^2, \; \forall \pr{k} \in \bar{\mc{U}}_{b_k}, \; \forall n \in \mc{C} \IEEEyessubnumber \label{eqn-dual-1e} \\
&\quad& \mbfa{\nu}^{\set{b_k}} \in \mathbb{R}^{n_{b_k}}_+, \; \eqref{eqn-8} \; \text{and} \; \eqref{eqn-primal-1c}. \IEEEyessubnumber \label{eqn-dual-1f}
\end{IEEEeqnarray}
The global interference variable \me{\mbfa{\zeta}(b_k)} in \eqref{eqn-dual-2a} can be dropped without affecting the optimal solution, since it is fixed for a given iteration. Instead of solving the dual decomposition problem defined in \eqref{eqn-dual-2}, we consider the \ac{ADMM} scheme which has better convergence properties as discussed in \cite{bertsekas1999nonlinear,boyd2011distributed}.

In order to limit the local interference assumptions \me{\zeta^{\set{b_k}}_{l,k,n,b}} at the \acp{BS} \me{b_k}, the \ac{ADMM} method augments a scaled quadratic penalty of the interference deviation between the local and consensus value for the interference from the \ac{BS} \me{b} as \me{\zeta_{l,k,n,b}} to the objective function. At optimality, the locally assumed and the consensus interference values will be equal,  providing no contribution to the objective function. The optimal step size use to update the dual variables is the scaling factor \me{\rho} used to scale the penalty term in the objective function \cite{boyd2011distributed}. Now, the subproblem at each \ac{BS} for the \me{\ith{i}} iteration is give by
\begin{IEEEeqnarray}{rCl} \label{eqn-dual-3}
\underset{\substack{\gamma_{l,k,n}, \mvec{m}{l,k,n}, \\ \beta_{l,k,n}, \zeta^{\set{b_k}(i)}_{l,k,n,b}}}{\text{minimize}} &\quad & \| \tilde{\mbf{v}}_{b_k} \|_q + \mbfa{\nu}^{{\set{b_k}(i-1)}T} \mbfa{\zeta}^{\set{b_k}(i)} + \frac{\rho}{2} \Big \| \, \underbrace{\mbfa{\zeta}^{(i-1)}(b_k)}_{\text{consensus values}} - \underbrace{\mbfa{\zeta}^{\set{b_k}(i)}}_{\text{local values}} \, \Big \|^2_2 \IEEEyessubnumber \label{eqn-dual-3a} \\
\text{subject to} & \quad & \eqref{eqn-dual-1d} - \eqref{eqn-dual-1f} \IEEEyessubnumber \label{eqn-dual-3b},
\end{IEEEeqnarray}
where the superscript \me{i} represents the current iteration or the information exchange index and \me{\mbfa{\zeta}^{(i-1)}} denotes the updated global interference level from the \me{\ith{(i-1)}} information exchange of the local interference vector \me{\mbfa{\zeta}^{\set{b}(i-1)}, \forall b \in \mc{B}}.

Once the local problems are solved at each \ac{BS}, the new update for the global interference vector \me{\mbfa{\zeta}^{(i)}} and the dual variables \me{\mbfa{\nu}^{\set{b}(i)}} are performed at each \ac{BS} independently by exchanging the corresponding local copies of the interference vector \me{\mbfa{\zeta}^{\set{b}(i)}, \forall b \in \mc{B}}. Since the entries in \me{\mbfa{\zeta}^{(i)}} relates exactly two \acp{BS} only, each entry in the \me{\mbfa{\zeta}^{(i)}} can be updated by exchanging the local copies from the corresponding two \acp{BS} only. For instance, the entry \me{\zeta^{(i)}_{l,\mc{U}_{b_k}(1),n,b}} depends on the local interference value \me{\zeta^{\set{b_k}(i)}_{l,\mc{U}_{b_k}(1),n,b}} assumed by the \ac{BS} \me{b_k} and the actual interference caused by the \ac{BS} \me{b} as in \me{\zeta^{\set{b}(i)}_{l,\mc{U}_{b_k}(1),n,b}} as
\begin{equation}
\zeta_{l,\mc{U}_{b_k}(1),n,b}^{(i)} = \frac{1}{2} \, \left ( \, \zeta^{\set{b}(i)}_{l,\mc{U}_{b_k}(1),n,b} + \zeta^{\set{b_k}(i)}_{l,\mc{U}_{b_k}(1),n,b} \, \right ).
\label{zeta_update}
\end{equation}
The dual variable entries in the vector \me{\mbfa{\nu}^{\set{b_k}}}, which is the stacked dual variables corresponding to the interference equality constraint at the \ac{BS} \me{b_k}, are updated using the subgradient as
\begin{equation}\label{dual-sg-update}
\nu^{\set{b_k}(i)}_{l,k,n,b} = \nu^{\set{b_k}(i-1)}_{l,k,n,b} + \rho \, \left (\zeta^{\set{b_k}(i)}_{l,k,n,b} - \zeta^{(i)}_{l,k,n,b} \right ), \forall b,b_k \in \mc{B}, \forall k \in \bar{\mc{U}}_{b}.
\end{equation}
